Tampering Detection In Low-Power Smart Cameras

### 2015-05-26 paper.tex
Messo immagine indicatori

%\newcommand{\gi}[1]{}
%\newcommand{\ad}[1]{}
%\newcommand{\cl}[1]{}

@@

\gi{Qui ST pu\'o aggiungere qualche esempio di applicazione qui o menzionare dispositivi di riferimento (con tanto di link a datasheet). Magari aggiungendo che le batterie recenti permettono operativit\'a di due anni a questi regimi.}\cl{Verifico cosa possiamo dire del SecSoc.}
%
In this work we expressly target low-power and ultra-low-power smart cameras, devices that operate at very low frame-rates (e.g. possibly less than one frame every minute), that are characterized by a constrained computational power and memory, and that are battery powered (typically lasting one or two years). As an example, consider Wireless Multimedia Sensor Networks (WMSN) \cite{akyildiz2007survey} where nodes are wirelessy connected smart-cameras, that can acquire and transmit images at regular time interval or upon requests. Often, in WMSN, the units are not connected to the power supply and have to operate with batteries and possibly rely on energy harvesting mechanisms~\gi{REFERENCES, Adriano, guarda se c'e' qualche reference dal nostro paper di TIM} \cite{magno2009adaptive}. Therefore, the power consumption is a serious concern in these devices.

@@

%What are the key components of my approach and results? Also include any specific limitations.
We present an algorithm to detect both camera defocus and displacement, as stated in Section \ref{sec:probForm}. The algorithm presented in Section~\ref{sec:propSol} relies on two indicators: the average image intensity to detect displacement, and the average gradient norm, to detect blur. These indicators can be easily computed and monitored by an outlier-detection technique to detect the respective tampering. We show that displacement-detection performance can be improved by means of a preliminarily partitioning of the scene into nonoverlapping regions, and then separately computing and monitoring the average intensity over different regions (Section \ref{subsec:Segmentation}). Remarkably, these regions need to be defined during an initial configuration phase of the smart camera, thus there is no relevant computational overhead w.r.t. monitoring the whole image. Experiments in Section \ref{sec:experiments} show that operating on image regions can substantially improve the detection of camera displacements. Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.

@@

\subsection{Related Works}\label{subsec:relWorks}
% Why hasn't it been solved before? (Or, what's wrong with previous proposed solutions? How does mine differ?)
The literature concerning tampering detection is mostly focused on video surveillance applications and operates at few frames per second \gi{Adriano, \'e vero?}. Background models are typically leveraged to identify defocus and occlusions; in particular  ~\cite{aksay2007camera} performs detect defocus \ad{defocus detection?} by analyzing the wavelet domain of each frame and performs histogram comparison for detecting occlusions, while camera displacements are not considered. A background-subtraction technique is employed in~\cite{saglam2009real} to identify defocus, occlusions, and displacements. \gi{FIXME Comparison in the Fourier domain for defocus detection, histogram comparison for occlusion detection, comparison between current background and delayed background for displacement detection}. Background subtraction is also used in \cite{gil2007automatic} to detect defocus, occlusions, and displacements. \gi{FIXME Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection}.
In contrast, no background models are used in \cite{alippi2010detecting}, where a sequential monitoring scheme based on a change-detection test is employed to detect changes in the average gradient energy of each frame to detect defocus due to external disturbances on the camera lens.  
\cite{ribnick2006real}: comparison between frames belonging to a buffer in order to find high values of dissimilarity, associated to tampering.
\cite{kryjak2012fpga}: implementation in a FPGA of a solution based on background modeling, histograms comparisons, edges comparisons.
\cite{harasse2004automated}: tampering detection inside a moving vehicle; uses background subtraction methods in order to identify defocus, occlusions, and displacements. Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection.
\cite{tsesmelis2013tamper}: monitoring of the number of key points extracted by SURF in order to detect defocus events, partition in blocks and HOG descriptors matching for each block in order to detect occlusions. These types of solutions requires a lot of computations
%
%
%
\section{Problem Formulation}\label{sec:probForm}
%
Let $z_t$ be the frame acquired at time $t$, which we model as
\begin{equation}
\label{eq:observationModel}
z_t(x)=\mathcal{D}_t[y_t](x) \quad \forall x \in X
\end{equation}
where $\mathcal{D}_t$ denotes the degradation operator that transforms the original image $y_t$ in the frame $z_t$; $X \subset \mathbb{Z}^2$ denotes the regular pixels grid and $x\in \mathbb{Z}^2$ indicates the pixel coordinates. As far as there is no tampering attacks/events,
\begin{equation}
\label{eq:no_tampering}
\mathcal{D}_t[y_t](x) = y_t(x) + \eta_t(x) \quad \forall x \in X
\end{equation}
where $\eta_t$ is a random variable accounting for different noise sources (e.g., thermal, quantization, photon-counting). In normal conditions, all the images $y_t$ (thus also the frames $z_t$) might show different content but are acquired from the same viewpoint and the same camera orientation.

When at time $\tau^*$ an external disturbance introduces \emph{blur/defocus}, the image $y_t$ is degraded by a spatially variant blur operator, and $z_t$ becomes
\begin{equation}
\label{eq:model_defocus}
\mathcal{D}_t[y_t](x) = \int_{\mathcal{X}}y(s)h_t(x,s)ds\, + \eta_t(x) \quad \forall x \in X, t \geq \tau^*
\end{equation}
where $h_t(x,\cdot) > 0$ is the point-spread function at pixel $x \in X$.

A \emph{camera displacement} at frame $\tau^*$ is instead modeled as 
\begin{equation}
\label{eq:model_displacement}
z_t(x)  = \left\{ \begin{array}{rcl}
y_t(x) + \eta(x) & \mbox{per} & t < T^* \\
w_t(x) + \eta(x) & \mbox{per} & t \geqslant T^*
\end{array}\right. ,
\end{equation}
where $w_t$ relates to a different viewpoint and/or camera orientation than $y_t$.

The proposed tampering-detection algorithm analyzes a sequence of frames $\{z_t\}_t$ to detect time $\tau^*$ when any tampering like \eqref{eq:model_defocus} or \eqref{eq:model_displacement} occurs. We require that the initial $T_0$ frames are without tampering and provided for training.

@@

Algorithm \ref{alg:DISPL} presents the proposed tampering detection, which relies on two indicators that are very simple to compute: the average intensity (also referred to as \emph{luma}, and denoted by $l$) and the average norm of the gradient (denoted by $g$). The former is meant to detect camera displacement, the latter camera defocus. As anticipated in Section \ref{sec:introduction}, during the initial configuration, the scene is segmented in $\{R_k\}, k = 1,\dots,K$ disjoint regions ($R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$). Details about the employed segmentation algorithm are in Section \ref{subsec:Segmentation}.

In each frame $z_t$ the luma is separately computed in of the $K$ regions,
\begin{equation}\label{eq:lumaRegions}
l_k(t) =\frac{1}{\#R_k} \sum_{R_k} z_t(x), \ \ k = 1, \dots, K\,,
\end{equation}
while the average gradient norm is computed over the whole image
\begin{equation}
\label{eq:normaGradiente}
g(t) = \sum_X \left (\sqrt{\left(z_t \circledast f_h\right)^2(x) + \left(z_t \circledast f_v\right)^2(x)}\right),
\end{equation}
where $f_h$ and $f_h$ are the horizontal and vertical derivative filters, respectively, $\#(\cdot)$ indicates the cardinality of a set and $\circledast$ the 2d convolution.

As shown in Figure REFERENCE\_TODO.a), tampering results in outliers in the sequence of indicators. Detecting these outliers is not straightforward, since indicators heavily fluctuates in normal conditions because of short-term variations in the depicted scene or illumination changes. For outlier-detection purposes~\cite{Chandola2009}, it is more convenient to operate on sequences of independent and identically distributed (i.i.d.) values, where outliers can be detected by density-based methods or simply by confidence regions that is what we considered in our algorithm. Unfortunately, trends or deterministic variations in the indicator sequences are not easy to compensate because of unpredictable evolutions of the scene.

To obtain indicators that in normal conditions can be treated as i.i.d. realizations of a random variable we detrend~\cite{Gustafsson2000} the indicator sequence by performing temporal derivative
\begin{equation}\label{eq:detrending}
 \partial l_k(t) = l_k(t)-l_k(t-1),  \ \ k = 1, \dots, K\,,
\end{equation}
and we similarly define $\partial g(t)$. Detrended indicators are more suitable to be monitored by a confidence interval since, as shown in Figure REFERENCE\_TODO.b). However, a tampering yielding permanent changes in the indicators, as those in Figure REFERENCE\_TODO.a results in a single outlier in the detrended indicator~\eqref{eq:detrending}.

\begin{algorithm}[t]
	% \SetAlgoNoLine
	\LinesNumbered
	\SetAlgoNlRelativeSize{0}
	\SetNlSty{small}{}{.}
	\textbf{Input}: $\gamma$, $T_{o}$, $\{R_k\} k=1,\dots,K$ \\
	\textbf{Configuration}:\\
	\lnl{DISPL-Tr1} Compute $\partial l_k(t)$ and $\partial g(t), t = 1, \cdots, T_o$\\ 
	\lnl{DISPL-Tr7} Compute $\sigma_{l_k}$ and $\sigma_{g}$, the standard deviations of $\partial l_k$ and $\partial g$.\\

	\textbf{Operational phase}:\\
	\lnl{DISPL-Test1} \For{$t=T_{o} + 1,\dots,\infty$}{
		\lnl{DISPL-Test2} Get frame $z_t$, set $n_l =0$
		\lnl{DISPL-Test4} \For{$k=1,\dots,K$}{
			\lnl{DISPL-Test5} Compute $\partial l_k(t)$\\
			\lnl{DISPL-Test6} \If{$\partial l_k(t) < -\gamma \sigma_{l_k} \vee \partial l_k(t) > \gamma \sigma_{l_k} $}{
				\lnl{DISPL-Test7} $n_l=n_l+1$\\
			}
		}
		\lnl{DISPL-Test8} \If{$n_l\geq K-1$}{
			\lnl{DISPL-Test9} tampering (probably displacement or occlusion) has occurred in $z_t$ \\
		}
		\lnl{DISPL-Test10} Compute $\partial g(t)$\\
		\lnl{DISPL-Test11} \If{$\partial g(t) < -\gamma \sigma_{g} \vee \partial l(t) > \gamma \sigma_{g} $}{
				\lnl{DISPL-Test9} tampering (probably out-of-focus) has occurred in $z_t$ \\
								}

	}   
	\caption{The Proposed Tampering-Detection Algorithm}
	\label{alg:DISPL}
\end{algorithm}

\subsection{Scene Segmentation}\label{subsec:Segmentation}
Consideriamo la sequenza $\{z_t\}$ di frame acquisiti dalla camera, con $t=1,\dots,T_{c}$.
Per ciascun pixel $x\in\mathcal{X}$ calcoliamo un vettore $\textbf{d}(x)$ di $5$ elementi
\begin{equation}
\label{eq:featureVector}
\textbf{d}(x)=\left[r(x);c(x);\mu_{\nabla}(x);\sigma_{\nabla}(x);\bar{z}(x)\right], \textbf{d}(x) \in \mathbb{R}^5
\end{equation}
dove:
\begin{itemize}
	\item $r(x)$ rappresenta il numero di riga del pixel $x$.
	\item $c(x)$ rappresenta il numero di colonna del pixel $x$.
	\item $\mu_{\nabla}(x)$ rappresenta il valore del gradiente nel pixel $x$ mediato nel tempo:
	\begin{equation}
	\label{eq:segmentazioneGrad}
	\mu_{\nabla}(x) = \frac{\sum_{t=1}^{T_c}(\|\nabla z_t\|_2^2 \circledast f)(x)}{T_c},
	\end{equation}
	dove abbiamo indicato con $\|\nabla z_t\|_2^2$ la norma del gradiente per l'immagine $z_t$, definita in \eqref{eq:normaGradiente}, e con $f$ il filtro gaussiano discreto derivato dal campionamento di \eqref{eq:gaussian}.
	\item $\sigma_{\nabla}(x)$ rappresenta la deviazione standard nel tempo del gradiente nel pixel $x$:
	\begin{equation}
	\label{eq:segmentazioneVar}
	\sigma_{\nabla}(x)=\sqrt{\frac{1}{T_c - 1}\sum_{t=1}^{T_c}\left(\left(\|\nabla z_t\|_2^2 \circledast f\right)(x)-\mu_{\nabla}(x)\right)^2}.
	\end{equation}
	\item $\bar{z}(x)$ rappresenta il valore della luma del pixel $x$ mediato nel tempo:
	\begin{equation}
	\label{eq:segmentazioneLuma}
	\bar{z}(x)=\frac{\sum_{t=1}^{T_c}( z_t \circledast f)(x)}{T_c}.
	\end{equation}
\end{itemize}

@@

\section{Experiments}\label{sec:experiments}
The proposed algorithm has been implemented in MATLAB, and has been tested on two datasets.
The first one refers to frame sequences taken from webcams recording some parts of cities (as in Figure \ref{fig:sequences}(a)), where we have introduced some synthetically generated tampering events:
defocus has been simulated using gaussian filters (\ref{fig:sequences}(b)), while displacement has been created with concatenation of similar frame sequences (Figure \ref{fig:sequences}(c)).
In some cases these sequences contained real tampering events, as in Figure \ref{fig:sequences}(d).\\
\begin{figure}[t]
\centering
\includegraphics[width=1\linewidth]{Immagini/sequenze}
\caption{Sequences taken from webcams: \textbf{(a)} no tampering events; \textbf{(b)} defocus event on 4-th and 5-th frames, created using gaussian filtering; \textbf{(c)} displacement event on 4-th and 5-th frames, created using concatenation between similar sequences; \textbf{(d)} real displacement event on 4-th frame.}
\label{fig:sequences}
\end{figure}
The second dataset was taken using a Raspberry Pi Model B+, with its camera module, and the tampering was introduced by moving the device or putting water on the camera.\\
Four figures of merit have been suggested to assess the performance of the proposed algorithm:\\
\begin{description}
	\item[TP]  True positives. It measures the number of tampering events correctly detected by the algorithm.\\ 
	\item[TN]  True negatives. It measures the number of frames without tampering that are not detected by the algorithm. \\ 
	\item[FP]  False positives. It measures the number of tampering events erroneously detected by the algorithm.\\ 
	\item[FN]  False negatives. It measures the number of tampering events not detected by the algorithm. \\ 
\end{description} 
These indicators are computed varying the parameter $\gamma$, which defines the thresholds for the one-shot monitoring. 
This permits us to generate \textit{ROC curves}, where on the x-axis there is:
\[1-\text{SPECIFICITY}_\gamma = 1-\frac{\text{TN}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma}=\frac{\text{FP}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma},\]
while on th y-axis there is:
\[\text{RECALL}_\gamma=\frac{\text{TP}_\gamma}{\text{TP}_\gamma+\text{FN}_\gamma}.\]
The construction of these curves permits us to make a comparison with respect to other methods.
In particular the alternative approaches that we have considered working:
\begin{itemize}
\item considering the whole scene for the features computation;
\item considering adaptive region, as described in Section \ref{sec:propSol}, for the feature computation;
\item considering voronoi regions \cite{aurenhammer1991voronoi}, which are easier to compute with respect to our solution but don't consider the scene content, for the feature computation.
\end{itemize}

>>>>

@@

\gi{Qui ST pu\'o aggiungere qualche esempio di applicazione qui o menzionare dispositivi di riferimento (con tanto di link a datasheet). Magari aggiungendo che le batterie recenti permettono operativit\'a di due anni a questi regimi.}\cl{Verifico cosa possiamo dire del SecSoc.}
%
In this work we expressly target low-power and ultra-low-power smart cameras, devices that operate at very low frame-rates (e.g. possibly less than one frame every minute), that are characterized by a constrained computational power and memory, and that are battery powered (typically lasting one or two years). As an example, consider Wireless Multimedia Sensor Networks (WMSN) \cite{akyildiz2007survey} where nodes are wirelessy connected smart-cameras, that can acquire and transmit images at regular time interval or upon requests. Often, in WMSN, the units are not connected to the power supply and have to operate with batteries and possibly rely on energy harvesting mechanisms~\gi{REFERENCES, Adriano, guarda se c'e' qualche reference dal nostro paper di TIM}. Therefore, the power consumption is a serious concern in these devices.

@@

%What are the key components of my approach and results? Also include any specific limitations.
We present an algorithm to detect both camera defocus and displacement, as stated in Section \ref{sec:probForm}. The algorithm presented in Section~\ref{sec:Algorithm} relies on two indicators: the average image intensity to detect displacement, and the average gradient norm, to detect blur. These indicators can be easily computed and monitored by an outlier-detection technique to detect the respective tampering. We show that displacement-detection performance can be improved by means of a preliminarily partitioning of the scene into nonoverlapping regions, and then separately computing and monitoring the average intensity over different regions (Section \ref{subsec:Segmentation}). Remarkably, these regions need to be defined during an initial configuration phase of the smart camera, thus there is no relevant computational overhead w.r.t. monitoring the whole image. Experiments in Section \ref{sec:experiments} show that operating on image regions can substantially improve the detection of camera displacements. Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.

@@

\subsection{Related Works}\label{subsec:relWorks}
% Why hasn't it been solved before? (Or, what's wrong with previous proposed solutions? How does mine differ?)
The literature concerning tampering detection is mostly focused on video surveillance applications and operates at few frames per second \gi{Adriano, \'e vero?}. Background models are typically leveraged to identify defocus and occlusions; in particular  ~\cite{aksay2007camera} performs detect defocus by analyzing the wavelet domain of each frame and performs histogram comparison for detecting occlusions, while camera displacements are not considered. A background-subtraction technique is employed in~\cite{saglam2009real} to identify defocus, occlusions, and displacements. \gi{FIXME Comparison in the Fourier domain for defocus detection, histogram comparison for occlusion detection, comparison between current background and delayed background for displacement detection}. Background subtraction is also used in \cite{gil2007automatic} to detect defocus, occlusions, and displacements. \gi{FIXME Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection}.
In contrast, no background models are used in \cite{alippi2010detecting}, where a sequential monitoring scheme based on a change-detection test is employed to detect changes in the average gradient energy of each frame to detect defocus due to external disturbances on the camera lens.  
\cite{ribnick2006real}: comparison between frames belonging to a buffer in order to find high values of dissimilarity, associated to tampering.
\cite{kryjak2012fpga}: implementation in a FPGA of a solution based on background modeling, histograms comparisons, edges comparisons.
\cite{harasse2004automated}: tampering detection inside a moving vehicle; uses background subtraction methods in order to identify defocus, occlusions, and displacements. Comparison of edges pixels count for defocus detection, entropy comparison for occlusion detection, block matching algorithm for displacement detection.
\cite{tsesmelis2013tamper}: monitoring of the number of key points extracted by SURF in order to detect defocus events, partition in blocks and HOG descriptors matching for each block in order to detect occlusions. These types of solutions requires a lot of computations
%
%
%
\section{Problem Formulation}\label{sec:probForm}
%
Let $z_t$ be the frame acquired at time $t$, which we model as
\begin{equation}
\label{eq:observationModel}
z_t(x)=\mathcal{D}_t[y_t](x) \quad \forall x \in X
\end{equation}
where $\mathcal{D}_t$ denotes the degradation operator that transforms the original image $y_t$ in the frame $z_t$; $X \subset \mathbb{Z}^2$ denotes the regular pixels grid and $x\in \mathbb{Z}^2$ indicates the pixel coordinates. As far as there is no tampering attacks/events,
\begin{equation}
\label{eq:no_tampering}
\mathcal{D}_t[y_t](x) = y_t(x) + \eta_t(x) \quad \forall x \in X
\end{equation}
where $\eta_t$ is a random variable accounting for different noise sources (e.g., thermal, quantization, photon-counting). In normal conditions, all the images $y_t$ (thus also the frames $z_t$) might show different content but are acquired from the same viewpoint and the same camera orientation.

When at time $\tau^*$ an external disturbance introduces \emph{blur/defocus}, the image $y_t$ is degraded by a spatially variant blur operator, and $z_t$ becomes
\begin{equation}
\label{eq:model_defocus}
\mathcal{D}_t[y_t](x) = \int_{\mathcal{X}}y(s)h_t(x,s)ds\, + \eta_t(x) \quad \forall x \in X, t \geq \tau^*
\end{equation}
%where  $\mathcal{B}_t$ corresponds to a linear operator
%\begin{equation}
%\label{eq:blur}
%\mathcal{B}[y](x) = \int_{\mathcal{X}}y(s)h(x,s)ds\,,
%\end{equation}
where $h_t(x,\cdot) > 0$ is the point-spread function at pixel $x \in X$.
%\begin{equation}
%\label{eq:blur_convolution}
%\mathcal{B}[y](x) = \int_{\mathcal{X}}y(s)h(s-x)ds,
%\end{equation}
%
%\begin{equation}
%\label{eq:blur_multi}
%z_t(x)=\mathcal{D}_t[y_t](x) = \mathcal{B}_t[y_t](x) + \eta(x), \qquad x \in \mathcal{X}.
%\end{equation}
A \emph{camera displacement} at frame $\tau^*$ is instead modeled as 
\begin{equation}
\label{eq:model_displacement}
z_t(x)  = \left\{ \begin{array}{rcl}
y_t(x) + \eta(x) & \mbox{per} & t < T^* \\
w_t(x) + \eta(x) & \mbox{per} & t \geqslant T^*
\end{array}\right. ,
\end{equation}
where $w_t$ relates to a different viewpoint and/or camera orientation than $y_t$.

The proposed tampering-detection algorithm analyzes a sequence of frames $\{z_t\}_t$ to detect time $\tau^*$ when any tampering like \eqref{eq:model_defocus} or \eqref{eq:model_displacement} occurs.

%Per $t<T^*$ abbiamo che i frame vengono acquisiti in condizioni di funzionamento normale:
%\[ z_t(x)=y_t(x) + \eta(x), \forall x \in \mathcal{X} \mbox{, per } t=1,\dots , T^*-1. \] 
%All'istante di tempo $t = T^*$ avviene un evento di tampering, il quale compromette i frame per $t\geq T^*$.  
%In particolare, nel caso di una sfocatura avremo
%\[ z_t = \mathcal{B}_t[y_t](x) + \eta(x), \forall x \in \mathcal{X} \mbox{, per } t \geq T^*,\]
%mentre nel caso di uno spostamento della camera avremo
%\[ z_t = w_t(x) + \eta(x), \forall x \in \mathcal{X} \mbox{, per } t \geq T^*. \]

@@

The proposed tampering-detection algorithm relies on two indicators that are very easy to compute: the average intensity (also refer to \emph{luma} and denoted by $l$) and the average norm of the gradient (denoted by $g$). To analyze an incoming frame, we compute these indicators at first and then detect possible tampering attacks by determining whether these falls outside a confidence interval defined during an initial training phase.

\begin{algorithm}[t]
	% \SetAlgoNoLine
	\LinesNumbered
	\SetAlgoNlRelativeSize{0}
	\SetNlSty{small}{}{.}
	\textbf{Input}: $\gamma$, $T_{o}$, $\{R_k\}, k=1,\dots,K$ \\
	\textbf{Configuration}:\\
%	\lnl{DISPL-Tr1} Extract regions $\{\mathcal{R}_k\}, k=1,\dots,K$  \\
	\lnl{DISPL-Tr2} \For{$t=1,\dots,T_{o}$}
	{	\lnl{DISPL-Tr3} Get frame $z_t$ \\
		\lnl{DISPL-Tr4} \For{$k=1,\dots,K$}{
			\lnl{DISPL-Tr5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $R_k$\\
		}
	}
	\lnl{DISPL-Tr6} \For{$k=1,\dots,K$}{
		\lnl{DISPL-Tr7} Compute $\sigma_l^k$\\
	}
	\textbf{Operational phase}:\\
	\lnl{DISPL-Test1} \For{$t=T_{o},\dots,\infty$}{
		\lnl{DISPL-Test2} Get frame $z_t$ \\
		\lnl{DISPL-Test3} $n_l = 0$ \\
		\lnl{DISPL-Test4} \For{$k=1,\dots,K$}{
			\lnl{DISPL-Test5} Compute $l^k(t)$, $\partial l^k(t)$ for the region $R_k$\\
			\lnl{DISPL-Test6} \If{$\partial l^k(t) < -\gamma \sigma_l^k \vee \partial l^k(t) > \gamma \sigma_l^k $}{
				\lnl{DISPL-Test7} $n_l=n_l+1$\\
			}
		}
		\lnl{DISPL-Test8} \If{$n_l\geq K-1$}{
			\lnl{DISPL-Test9} $z_t$ is a tampered frame\\
		}
	}   
	\caption{Tampering detection algorithm}
	\label{alg:DISPL}
\end{algorithm}

\subsection{Scene Segmentation}\label{subsec:Segmentation}
Consideriamo la sequenza $\{z_t\}$ di frame acquisiti dalla camera, con $t=1,\dots,T_{c}$.
Per ciascun pixel $x\in\mathcal{X}$ calcoliamo un vettore $\textbf{d}(x)$ di $5$ elementi
\begin{equation}
\label{eq:featureVector}
\textbf{d}(x)=\left[r(x);c(x);\mu_{\nabla}(x);\sigma_{\nabla}(x);\bar{z}(x)\right], \textbf{d}(x) \in \mathbb{R}^5
\end{equation}
dove:
\begin{itemize}
	\item $r(x)$ rappresenta il numero di riga del pixel $x$.
	\item $c(x)$ rappresenta il numero di colonna del pixel $x$.
	\item $\mu_{\nabla}(x)$ rappresenta il valore del gradiente nel pixel $x$ mediato nel tempo:
	\begin{equation}
	\label{eq:segmentazioneGrad}
	\mu_{\nabla}(x) = \frac{\sum_{t=1}^{T_c}(\|\nabla z_t\|_2^2 \circledast f)(x)}{T_c},
	\end{equation}
	dove abbiamo indicato con $\|\nabla z_t\|_2^2$ la norma del gradiente per l'immagine $z_t$, definita in \eqref{eq:normaGradiente}, e con $f$ il filtro gaussiano discreto derivato dal campionamento di \eqref{eq:gaussian}.
	\item $\sigma_{\nabla}(x)$ rappresenta la deviazione standard nel tempo del gradiente nel pixel $x$:
	\begin{equation}
	\label{eq:segmentazioneVar}
	\sigma_{\nabla}(x)=\sqrt{\frac{1}{T_c - 1}\sum_{t=1}^{T_c}\left(\left(\|\nabla z_t\|_2^2 \circledast f\right)(x)-\mu_{\nabla}(x)\right)^2}.
	\end{equation}
	\item $\bar{z}(x)$ rappresenta il valore della luma del pixel $x$ mediato nel tempo:
	\begin{equation}
	\label{eq:segmentazioneLuma}
	\bar{z}(x)=\frac{\sum_{t=1}^{T_c}( z_t \circledast f)(x)}{T_c}.
	\end{equation}
\end{itemize}
%
%
%
\subsection{Indicators}\label{subsec:Indicators}

\begin{equation}
\label{eq:gradientRegions}
\begin{array}{ccc}
g^k(t)&  = & \mathcal{G}^k[z_t] = \frac{\sum_{R_k}\| \nabla z_t(x) \| _2^2 }{|{R_k}|}\\
\partial g^k(t) & =& g^k(t)-g^k(t-1) 
\end{array},
\end{equation}

In particolare, per il calcolo delle derivate orizzontali  abbiamo utilizzato il seguente filtro $f_h$:
\[f_h = f \circledast \left[ \begin{array}{rcl}
1 & 0 & -1
\end{array}\right], \] 
mentre per il calcolo delle derivate verticali abbiamo utilizzato il seguente filtro $f_v$:
\[f_v = f \circledast \left[ \begin{array}{r}
1 \\ 0 \\ -1
\end{array}\right], \]
dove abbiamo indicato con $\circledast$ l'operatore di convoluzione.
Il filtro $f$, invece, \`e ottenuto tramite un campionamento della \textit{funzione gaussiana} $h$, con media $0$ e deviazione standard $\sigma$
\begin{equation}
\label{eq:gaussian}
h(i,j)=\frac{1}{2\pi\sigma^2}\exp\left(-\frac{i^2+j^2}{2\sigma^2}\right),
\end{equation}
e ponendo il valore massimo di questa funzione nel centro del filtro.
Con questi filtri \`e possibile calcolare la \textit{norma del gradiente} nel seguente modo:
\begin{equation}
\label{eq:normaGradiente}
\| \nabla z_t(x) \|_2^2=\left(z_t \circledast f_h\right)(x)^2 + \left(z_t \circledast f_v\right)(x)^2.
\end{equation}
Una volta calcolata la norma del gradiente \`e possibile farne la media come specificato in \eqref{eq:energyGradient}.
Il risultato finale \`e un indicatore \textit{scalare} per ciascun frame acquisito, che pu\`o essere monitorato per individuare eventi di sfocature. 
In particolare ci aspettiamo che l'evento di sfocatura provochi un abbattimento del valore di $g$.

\begin{equation}
\label{eq:lumaRegions}
\begin{array}{ccc}
l^k(t)&  = & \mathcal{L}^k[z_t] = \frac{\sum_{R_k} z_t(x) }{|{R_k}|}\\
\partial l^k(t) & =& l^k(t)-l^k(t-1) 
\end{array},
\end{equation}

@@

\section{Experiments}\label{sec:experiments}
The proposed algorithm has been implemented in MATLAB, and has been tested on two datasets.
The first one refers to frame sequences taken from webcams recording some parts of cities (as in Figure \ref{fig:sequences}(a)), where we have introduced some synthetically generated tampering events:
defocus has been simulated using gaussian filters (\ref{fig:sequences}(b)), while displacement has been created with concatenation of similar frame sequences (Figure \ref{fig:sequences}(c)).
In some cases these sequences contained real tampering events, as in Figure \ref{fig:sequences}(d).\\
\begin{figure}[t]
\centering
\includegraphics[width=1\linewidth]{Immagini/sequenze}
\caption{Sequences taken from webcams: \textbf{(a)} no tampering events; \textbf{(b)} defocus event on 4-th and 5-th frames, created using gaussian filtering; \textbf{(c)} displacement event on 4-th and 5-th frames, created using concatenation between similar sequences; \textbf{(d)} real displacement event on 4-th frame.}
\label{fig:sequences}
\end{figure}
The second dataset was taken using a Raspberry Pi Model B+, with its camera module, and the tampering was introduced by moving the device or putting water on the camera.\\
Four figures of merit have been suggested to assess the performance of the proposed algorithm:\\
\begin{description}
	\item[TP]  True positives. It measures the number of tampering events correctly detected by the algorithm.\\ 
	\item[TN]  True negatives. It measures the number of frames without tampering that are not detected by the algorithm. \\ 
	\item[FP]  False positives. It measures the number of tampering events erroneously detected by the algorithm.\\ 
	\item[FN]  False negatives. It measures the number of tampering events not detected by the algorithm. \\ 
\end{description} 
These indicators are computed varying the parameter $\gamma$, which defines the thresholds for the one-shot monitoring. 
This permits us to generate \textit{ROC curves}, where on the x-axis there is:
\[1-\text{SPECIFICITY}_\gamma = 1-\frac{\text{TN}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma}=\frac{\text{FP}_\gamma}{\text{TN}_\gamma+\text{FP}_\gamma},\]
while on th y-axis there is:
\[\text{RECALL}_\gamma=\frac{\text{TP}_\gamma}{\text{TP}_\gamma+\text{FN}_\gamma}.\]
The construction of these curves permits us to make a comparison with respect to other methods.
In particular the alternative approaches that we have considered working:
\begin{itemize}
\item considering the whole scene for the features computation;
\item considering adaptive region, as described in Section \ref{sec:propSol}, for the feature computation;
\item considering voronoi regions \cite{aurenhammer1991voronoi}, which are easier to compute with respect to our solution but don't consider the scene content, for the feature computation.
\end{itemize}
%
%% \subsection{Alternative Approaches}\label{subsec:AlternativeApproaches}
%
%
%\subsection{Performance Assessment}
%\gi{Adriano: Dire come vengono calcolate le ROC curves TPR e FPR, le cifre di merito insomma, spiegando bene che parametro varia}
%
%\gi{Adriano: metti entrambe le ROC curves, affiancate e per bene ed alcuni esempi di sequenze}
\begin{figure}[htb]
\centering
\subfigure[]{\label{fig:ROCdisplacement_luma}\includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement_luma}}
\subfigure[]{\label{fig:ROCdisplacement_fd} \includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement_fd}}
\caption{ROC curves for displacement detection, considering three alternative approaches. \textbf{(a)} Analysis of the mean luma energy. \textbf{(b)} Analysis of the frame differencing.}
\label{fig:ROCdisplacement}
\end{figure}
\begin{figure}[htb]
\centering
\subfigure[]{\label{fig:buenosAiresDef}\includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus1}}
\subfigure[]{\label{fig:buenosAiresDispl} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus_fd}}
\caption{ROC curves for defocus detection, considering three alternative approaches. \textbf{(a)} Analysis of the mean gradient energy. \textbf{(b)} Analysis of the frame differencing.}
\label{fig:ROCdisplacement}
\label{fig:ROCdefocus}
\end{figure}
Experimental results are shown in Figures \ref{fig:ROCdisplacement} and \ref{fig:ROCdefocus}.
As we could see there is an improvement, in the detection of camera displacements, with our approach which separately analizes the behavior of indicators in adaptive regions.
On the other side, as illustrated in Figure \ref{fig:ROCdefocus}, when monitoring indicators meant to detect blur/defocus, it is more convenient to consider the whole scene at once.
\subsection{Discussion}
\gi{Adriano: Aggiungi qua la complessit\'a computazionale}
% Please add the following required packages to your document preamble:
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
\begin{table}[tbh]
\centering
\begin{tabular}{l|l|l|l|l|l|l|}
\cline{2-7}
& \multicolumn{3}{l|}{\cellcolor[HTML]{C0C0C0}\textbf{Displacement Detection}}  & \multicolumn{3}{l|}{\cellcolor[HTML]{C0C0C0}\textbf{Defocus Detection}} \\ \cline{2-7} 
& \cellcolor[HTML]{EFEFEF}\textbf{Regions} & \cellcolor[HTML]{EFEFEF}\textbf{Total} & \cellcolor[HTML]{EFEFEF}\textbf{Voronoi} & \cellcolor[HTML]{EFEFEF}\textbf{Regions} & \cellcolor[HTML]{EFEFEF}\textbf{Total} & \cellcolor[HTML]{EFEFEF}\textbf{Voronoi} \\ \hline
\multicolumn{1}{|l|}{\cellcolor[HTML]{EFEFEF}\textbf{Luma}}     & 0.9994 & 0.9989 & 0.9985  &            &            &             \\ \hline
\multicolumn{1}{|l|}{\cellcolor[HTML]{EFEFEF}\textbf{Gradient}} &  		 &  		  &             & 0.9895 & 0.9996 & 0.9996  \\ \hline
\multicolumn{1}{|l|}{\cellcolor[HTML]{EFEFEF}\textbf{FD}}         & 0.9974 & 0.9944 & 0.9974  & 0.5526 & 0.5322 & 0.5391  \\ \hline
\end{tabular}
\caption{Area under curve of the analized solutions}
\label{tab:AUC}
\end{table}
